%!TEX root = paper.tex

Gas pump skimming is one of the most pervasive forms of payment card attacks in
	the U.S. today.
%
Criminals can easily bypass physical security (e.g. pump locks, cameras) at gas stations to install a skimmer on the internal payment circuitry of a gas pump.
%
Furthermore, many gas pump skimmers have Bluetooth access links; these allow criminals to exfiltrate stolen payment card data conveniently from the safety of their cars. 
%

For victims and investigators alike, these skimmers are hard to detect. 
%
They are hard to visually identify from outside with the pump door closed, and blend into the wiring even when the pump door is open.
%
These are also hard to identify in smartphone Bluetooth scans -- there are a lot of Bluetooth devices at gas stations (e.g. phones, fitness bands, OBD readers, car stereos, streetlights), and Bluetooth skimmers appear like any benign devices.
%
This
%Gas pump skimmers are easy to install : criminals can open gas pump enclosures
%	and hide a skimmer in internal payment wiring.
%
%Internal gas pump skimmers are also difficult to detect: they are not visible
%	from outside of the pump, and they blend into the wiring inside the pump.
%
%whAs a result, officials have resorted to detecting skimmers by performing
%	laborious manual inspections of the wiring inside gas pumps.
% 
%In addition, criminals can also avoid being caught using skimmers: many gas
%	pump skimmers have Bluetooth connectivity, allowing criminals to collect
%	payment data safely from inside their car.
%	These skimmers are 

In this chapter, I present a real-world empirical study aimed at understanding whether smartphone Bluetooth scans provide sufficient information to uniquely distinguish Bluetooth skimmers from the myriad of Bluetooth devices at a gas station.
%
In this 19-month measurement study, we collected Bluetooth scans at \visitedgasstations~gas stations across six states in the US, detecting \totalskimmers~Bluetooth-based skimmers among thousands of other Bluetooth devices at the stations. 
%

We observed that indeed Bluetooth links in skimmers can be differentiated from other legitimate devices using Bluetooth scan data.
%
However, legitimate devices may also utilize the same Bluetooth modules as skimmers and that is a common source of confusion.
%
In this study, we perform a comprehensive analysis of how su


