%!TEX root = paper.tex
\section{Future Work and Conclusion}
\label{sec:conclusion}

As new skimmer detection tools gain popularity, criminals will adapt skimming designs to evade detection.
%
We expect future skimmers will use techniques such those described in Section \ref{sec:hiding}.
%
Similar to Bluetana, future work in this area should emphasize designing easy-to-deploy systems for detecting
skimmers, and evaluating their effectiveness with large-scale studies.

Push-back from banks and card issuers has led to wide-scale adoption of EMV in
retail PoS systems. 
%
However, EMV adoption in gas stations across the U.S. has been slow due to high
costs.
%
Therefore, Visa and Mastercard have pushed the EMV adoption deadline for gas
stations from 2017 to October 2020~\cite{emv2020}. 
%
As gas stations begin migrating to EMV, skimmers targeting EMV will become more
common. 
%
Future research should focus on the detection of EMV ``shimmers'' that are
gaining in popularity.

Finally, we believe gas pump skimming is the harbinger of an era of attacks
using illicit wireless access links. 
%
For example, there is an internal Bluetooth-based implant for unlocking door
access control systems~\cite{blekey}. 
%
Future work should also identify other
systems that are vulnerable to using such illicit links.

In this chapter, we presented results of a 19-month-long measurement study of
Bluetooth scanning as a mechanism to detect illicit internal gas pump skimmers. 
%
Our evaluation showed that link layer characteristics of Bluetooth-based internal skimmers
can be distinguished from other Bluetooth devices commonly seen at gas
stations. 
%
We detected, and LE recovered, \totalskimmers~skimmers at 34 gas
stations across four states in the U.S. 
%
For \totalskimmersBluetana~of the detected skimmers, Bluetana was the only
source of information that prompted investigators to conduct an inspection.
%
In conclusion, link-layer information revealed in Bluetooth scans is effective at detection of illicit wireless access links at public locations, even in the presence of tens of other Bluetooth devices.

\begin{comment}
This paper presented Bluetana, a empirical measurement study of Bluetooth's ability to detect internal
card skimmers. Bluetooth makes it possible to quickly and accurately detect skimmers within gas dispensers.
Our evaluation has shown that Bluetooth is a promising method of skimmer detection, capable of
detecting ~\totalskimmers~ skimmers for a \emph{daily} monetary impact of ~\Bluetanafraudprevented~ dollars in a study of ~\visitedgasstations~ stations.
\end{comment}

%
\begin{comment}
However, gas stations have not adapted.
%
We randomly sampled 0.5\% of all gas stations in the US and classified pump payment technology.
%
Google Street View images of the dispensers revealed that as of the current date only 13.8\% of gas stations have
retrofitted EMV readers and only 1.6\% have installed EMV based dispensers.
\end{comment}
