\begin{vita}
    \noindent
    \begin{cv}{}
    \begin{cvlist}{}
    \item[2012] Bachelor of Engineering in Electronics and Instrumentation, \\
    Birla Institute of Technology and Science Pilani
    \item[2012--2015] Embedded Application Engineer, Texas Instruments Inc.
    \item[2015--2017] Teaching Assistant, Computer Science and Engineering\\University of California, San Diego
    \item[2017--2023] Research Assistant, University of California, San
    Diego
    \item[2019] Master of Science in Computer Science (Computer Engineering), \\
    University of California San Diego
    \item[2023] Doctor of Philosophy in Computer Science (Computer Engineering), \\
    University of California San Diego
    \end{cvlist}
    \end{cv}
    
    % This puts in the PUBLICATIONS header. Note that it appears inside
    % the vita environment. It is optional.
    \publications
    
    \noindent Nikoofard, A., Givehchian, H., Bhaskar, N., Schulman, A., Bharadia, D. and Mercier, P.P., 2022. Protecting Bluetooth User Privacy through Obfuscation of Carrier Frequency Offset. IEEE Transactions on Circuits and Systems II: Express Briefs.
    \\~\\
    \noindent Subbaraman, R., Bhaskar, N., Crow, S., Khazraee, M., Schulman, A. and Bharadia, D., 2022, June. Observing wideband RF Spectrum with low-cost, resource limited SDRs. In Proceedings of the 20th Annual International Conference on Mobile Systems, Applications and Services (pp. 617-618).
    \\~\\
    \noindent Givehchian, H., Bhaskar, N., Herrera, E.R., Soto, H.R.L., Dameff, C., Bharadia, D. and Schulman, A., 2022, May. Evaluating Physical-layer BLE Location Tracking Attacks on Mobile Devices. In 2022 IEEE Symposium on Security and Privacy (SP) (pp. 1690-1704). IEEE.
    \\~\\
    \noindent Jagtap, D., Bhaskar, N. and Pannuto, P., 2021, June. Century-scale Smart Infrastructure. In Proceedings of the Workshop on Hot Topics in Operating Systems (pp. 72-78).
    \\~\\
    \noindent Bhaskar, N., Bland, M., Levchenko, K. and Schulman, A., 2019, August. Please Pay Inside: Evaluating Bluetooth-based Detection of Gas Pump Skimmers. In USENIX Security Symposium (pp. 373-388).
    % This puts in the FIELDS OF STUDY. Also inside vita and also
    % optional.
\end{vita}