\documentclass[10pt]{article}
\usepackage{ucsdcseletterhead}
\usepackage{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{mathptmx}
\usepackage{courier}
\usepackage[scaled=.92]{helvet}
\usepackage[T1]{fontenc}
\pagestyle{empty}
\begin{document}
\hfill\today

James Antony, Dean\\
Graduate Division\\
UC San Diego

\vskip\baselineskip
Dear Dean Antony:

We request permission for Nishant Bhaskar to use the following publications
in his doctoral dissertation. Mr.\ Bhaskar was the principal
researcher/author on this paper.

Nishant Bhaskar, Maxwell Bland, Kirill Levchenko, and Aaron Schulman. ``Please Pay Inside: Evaluating Bluetooth-based Detection of Gas Pump Skimmers.`` In USENIX Security Symposium, pp. 373-388. 2019.

Hadi Givehchian, Nishant Bhaskar, Eliana Rodriguez Herrera, Héctor Rodrigo López Soto, Christian Dameff, Dinesh Bharadia, and Aaron Schulman. ``Evaluating Physical-Layer BLE location Tracking Attacks on Mobile Devices.`` In 2022 IEEE Symposium on Security and Privacy (SP), pp. 1690-1704. IEEE, 2022.

\vskip.5\baselineskip\centerline{$\sim$AND$\sim$}

We request permission for Nishant Bhaskar to use material currently being
prepared for submission for publication. Mr. Bhaskar was the principal
researcher/author on this paper.

Nishant Bhaskar, Raghav Subbaraman, Sam Crow, Moein Khazraee, Dinesh Bharadia, Aaron Schulman  ``HyperScanner: Fast Bluetooth Scanning``

\vskip3\baselineskip
\parbox[t]{.5\textwidth}{%
Professor Aaron Schulman\\
Chair of Committee}%
\parbox[t]{.5\textwidth}{%
Nishant Bhaskar}
\end{document}
